### 2015-01-05 to_do_list.tex
add copy of to_do_list.tex from pe_shocks parent

\documentclass[12pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-.7in} %left margin
\addtolength{\evensidemargin}{-.7in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{0.0in}
\setlength{\headsep}{0.0in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.5in}
\renewcommand{\baselinestretch}{1.0}
\setlength{\parindent}{0pt}
\linespread{1.1}

\usepackage{amssymb, amsmath, amsthm, bm}
\usepackage{graphicx,csquotes,verbatim}
\usepackage[backend=biber,block=space,style=authoryear]{biblatex}
\setlength{\bibitemsep}{\baselineskip}
\usepackage[american]{babel}
%dell laptop
\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
\renewcommand{\newunitpunct}{,}
\renewbibmacro{in:}{}

\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{xcolor}
\hbadness=10000

\newcommand{\ve}{\varepsilon}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\ta}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\Ta}{\Theta}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\btn}{\bm{\tau^n}}
\newcommand{\ga}{\gamma}
\newcommand{\Ga}{\Gamma}
\newcommand{\de}{\delta}

\begin{document}
\begin{center}
Ex-ante Lobbying
\end{center}
%\chead{to_do_list.tex}

\vskip.2in
\begin{itemize}
	\item Need to convey the possibilities of this set-up for capturing the real life dynamics of shocks integrated with lobbying dynamics, how that interacts with enforcement and administered protection
	\item Supply air tight argument / reference about $C+aW$ being equivalent to $\ga$ increasing everywhere
		\begin{itemize}
			\item CAN'T! It's not true! Only microfound a version with $1+a$ on those who lobby, $a$ on those who don't
		\end{itemize}
	\item Replace $e_{W,e}$ with $e^R_{W,e}$
	\item perhaps re-emphasize that gov't can't completely control lobbying: can't make it higher than lobby's optimal. But this is b/c there's only one lobby
	\item finish escape clause result
	\item Result 2: With ex-post lobbying only, it's possible to get no internalization of TOT externality when joint welfare is everywhere increasing in $\ga$. Can have trade agreement doing nothing.
		\begin{itemize}
			\item When TOT externality is internalized, welfare is higher at a \textit{given} $\ta$ because each country is not imposing the externality on the other
				\begin{itemize}
					\item But if $\ga$ is endogenous only, when they use the TA to maximize joint welfare, if joint welfare is everywhere increasing in $\ga$, they may not want to place a restriction on $\tau$ in relation to $\tau^N$
					\item In general, they may still do something that looks like internalizing TOT externality, but not in the case where joint welfare is everywhere increasing in $\ga$ (which is equivalent to $W$ increasing in $\ga$ in the separable, symmetric economy)
					\item In fact, the maximization problem looks exactly like unilateral maximization with a domestic commitment motive (reign in lobbying in second stage)
				\end{itemize}
		\end{itemize}
	\item pg. 24 calculations
		\begin{itemize}
			\item Decide where political objective fcn. section goes now
		\end{itemize}
	\item re-write intro
		\begin{itemize}
			\item MRC-style results survive incorporation of ex-ante lobbying. TA still works to restrict lobbying in many cases, but also to keep lobbyists ``in the game''
			\item Baldwin-style gov't welfare function works for a set of questions, but not for this one
				\begin{itemize}
					\item Need to be able to say which predictions/design questions are okay: those that don't change lobbying incentives
				\end{itemize}
			\item Coates and Ludema (2001)---endogenous (add to list of GH, MRC)
			\item How do I want to sell the paper? In relation to what literature? (what lit does endogenous? what lit doesn't?)
			\item What if the WTO actually \textit{did} what the literature says it does? It wouldn't work
			\item Be more explicit about relationship between $\ga(e)$ and $\ga(s)$
		\end{itemize}
\end{itemize}

\vskip.5in
For after this draft
\begin{itemize}
	\item Could escape clause be made to work with some kind of dynamic use constraint?
		\begin{itemize}
			\item When would lobby exert effort to top up?
		\end{itemize}
	\item What units are $\pi(\tau)$ and $e$ measured in? (no numeraire)
	\item existence proofs
\end{itemize}

\end{document}